---
title: "Activity Monitoring Prediction"
author: "Akash Suresh"
date: "August 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

### What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

### Reproducibility

Due to security concerns with the exchange of R code, your code will not be run during the evaluation by your classmates. Please be sure that if they download the repo, they will be able to view the compiled HTML version of your analysis.

## Exploratory Analysis

Lets load all the required packages and the data given.

```{r load,warning=FALSE,cache=FALSE,message=FALSE}
library(dplyr)
library(caret)
library(rpart)
library(randomForest)

## Replacing empty values with NA

training <- read.csv("C:/Users/Admin/Documents/Coursera/R/data/Course8/Activity-Monitoring-Prediction/data/pml-training.csv",na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv("C:/Users/Admin/Documents//Coursera/R/data/Course8/Activity-Monitoring-Prediction/data/pml-testing.csv",na.strings=c("NA","#DIV/0!", ""))
```

Looking the data,

```{r dim,warning=FALSE,message=FALSE}
dim(training)
```

There are 160 variables in the training set. On looking at the variables, few of the first ones seem to be extraneous. Namely,

```{r extra,warning=FALSE,message=FALSE}
names(training)[1:7]
```

Fields other than the user does not look to be required for the current model. Also, note that if timestamp is to be included in the set of predictors, make sure to remove the extraneous time attributes. X is a unique identifier that is not useful. Remove those attirbutes as it might contribute to overfitting. 

```{r filter,warning=FALSE,comment=FALSE}
training <- training[,-c(1,3:7)]
testing <- testing[,-c(1,3:7)]

set.seed(12321)

train <- createDataPartition(training$classe, p=0.8, list=FALSE)
modelTraining <- training[train, ] 
modelTesting <- training[-train, ]
```

## Models

Since, this problem requires a classification, we could try a decision tree first. However, this might give an overfitted model.

```{r decisionTree,warning=FALSE,message=FALSE}
modelTree <- rpart(classe ~ ., modelTraining, method="class")
predictionTree <- predict(modelTree, modelTesting, type = "class")
confusionMatrix(predictionTree, modelTesting$classe)$overall
```


The accuracy is pretty good as it is achieved with minimal data preprocessing. Lets see if Random Forests do any better.

```{r randomForest,warning=FALSE,message=FALSE,eval=FALSE}
modelForest <- randomForest(classe ~. , modelTraining, method = "class")
predictionForest <- predict(modelForest, modelTesting, type = "class")
confusionMatrix(predictionForest, modelTesting$classe)$overall
```

`
Overall Statistics

               Accuracy : 0.9947   
               
                 95% CI : (0.9922, 0.9965)
                 
    No Information Rate : 0.2845        
    
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.9933      
                  
 Mcnemar's Test P-Value : NA   
 
`

We clearly see that the accuracy with this model is much higher, almost a 100 percent. The predictions on the test data look like this.

```{r predictions,warning=FALSE,message=FALSE,eval=FALSE}
predictionForest <- predict(modelForest, testing, type = "class")
predictionForest
```

` 1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 19 20`

`B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B`

`Levels: A B C D E`


